High-level cognition during story listening is reflected in high-order dynamic correlations in neural activity patterns

### 2021-07-02 paper/main.tex

\documentclass[english]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{apacite}
\usepackage{hyperref}
\usepackage[sort]{natbib}
\usepackage{pxfonts}
\usepackage[utf8]{inputenc}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage[left]{lineno}
\usepackage{soul}
\linenumbers

@@

\begin{abstract}
  Our thoughts arise from coordinated patterns of interactions between
  brain structures that change with our ongoing experiences.
  High-order dynamic correlations in neural activity patterns reflect
  different subgraphs of the brain's functional connectome that display
  homologous lower-level dynamic correlations.  We tested the
  hypothesis that high-level cognition is reflected in high-order
  dynamic correlations in brain activity patterns.  We developed an
  approach to estimating high-order dynamic correlations in timeseries
  data, and we applied the approach to neuroimaging data collected as
  human participants either listened to a ten-minute story or listened
  to a temporally scrambled version of the story.  We trained
  across-participant pattern classifiers to decode (in held-out data)
  when in the session each neural activity snapshot was collected.  We
  found that classifiers trained to decode from high-order dynamic
  correlations yielded the best performance on data collected as
  participants listened to the (unscrambled) story.  By contrast,
  classifiers trained to decode data from scrambled versions of the
  story yielded the best performance when they were trained using
  first-order dynamic correlations or non-correlational activity
  patterns.  We suggest that as our thoughts become more complex, they
  are reflected in higher-order patterns of dynamic network
  interactions throughout the brain.
\end{abstract}

@@

\section*{Introduction}
A central goal in cognitive neuroscience is to elucidate the
\textit{neural code}: the mapping between (a) mental states or
cognitive representations and (b) neural activity patterns. One means
of testing models of the neural code is to ask how accurately that
model is able to ``translate'' neural activity patterns into known (or
hypothesized) mental states or cognitive
representations~\citep[e.g.,][]{HaxbEtal01, NormEtal06b, TongPrat12,
  MitcEtal08a, KamiTong05, NishEtal11, PereEtal18, HuthEtal12,
  HuthEtal16}.  Training decoding models on different types of neural
features (Fig.~\ref{fig:patterns}a) can also help to elucidate which
specific aspects of neural activity patterns are informative about
cognition and, by extension, which types of neural activity patterns
might compose the neural code.  For example, prior work has used
region of interest analyses to estimate the anatomical locations of
specific neural representations~\citep[e.g.,][]{EtzeEtal09}, or to
compare the relative contributions to the neural code of multivariate
activity patterns versus dynamic correlations between neural activity
patterns~\citep[e.g.,][]{MannEtal18, FongEtal19}.  An emerging theme
in this literature is that cognition is mediated by dynamic
interactions between brain structures~\citep{Gros88, Fris00, SporHone06, BassEtal06,
  Turk13, DemeEtal19, SoloEtal19, LuriEtal18, PretEtal17, ZouEtal19,
  MackEtal17, BresKels01, McIn00}.

@@

Studies of the neural code to date have primarily focused on
univariate or multivariate neural patterns~\citep[for review
see][]{NormEtal06b}, or (more recently) on patterns of dynamic
first-order correlations~\citep[i.e., interactions between pairs of
brain structures;][]{MannEtal18, FongEtal19, LuriEtal18, PretEtal17,
  ZouEtal19, DemeEtal19}.  What might the future of this line of work
hold?  For example, is the neural code implemented through
higher-order interactions between brain structures~\citep[e.g.,
see][]{ReimEtal17}?  Second-order correlations reflect
\textit{homologous} patterns of correlation.  In other words, if the
dynamic patterns of correlations between two regions, $A$ and $B$, are
similar to those between two other regions, $C$ and $D$, this would be
reflected in the second-order correlations between ($A$--$B$) and
($C$--$D$).  In this way, second-order correlations identify
similarities and differences between subgraphs of the brain's
connectome.  Analogously, third-order correlations reflect homologies
between second-order correlations-- i.e., homologous patterns of
homologous interactions between brain regions.  More generally,
higher-order correlations reflect homologies between patterns of
lower-order correlations.  We can then ask: which ``orders'' of
interaction are most reflective of high-level cognitive processes?

@@

To gain insights into the above aspects of the neural code, we
developed a computational framework for estimating dynamic high-order
correlations in timeseries data. This framework provides an important
advance, in that it enables us to examine patterns of higher-order
correlations that are computationally intractable to estimate via
conventional methods.  Given a multivariate timeseries, our framework
provides timepoint-by-timepoint estimates of the first-order
correlations, second-order correlations, and so on.  Our approach
combines a kernel-based method for computing dynamic correlations in
timeseries data with a dimensionality reduction step
(Fig.~\ref{fig:patterns}b) that projects the resulting dynamic
correlations into a low-dimensional space.  We explored two
dimensionality reduction approaches: principle components
analysis~\citep[PCA;][]{Pear01}, which preserves an approximately
invertible transformation back to the original data~\citep[e.g., this
follows related approaches taken by][]{McInJirs19, TokeSomm19,
  GonzEtal19}; and a second non-invertible algorithm for computing
dynamic patterns in eigenvector centrality~\citep{Land95}.  This
latter approach characterizes correlations between each feature
dimension's relative \textit{position} in the network (at each moment
in time) in favor of the specific activity histories of different
features~\citep[also see][]{BetzEtal19, SizeEtal18, ReimEtal17}.

@@

\section*{Results}
We sought to understand whether high-level cognition is reflected in
dynamic patterns of high-order correlations.  To that end, we
developed a computational framework for estimating the dynamics of
stimulus-driven high-order correlations in multivariate timeseries
data (see \textit{Dynamic inter-subject functional connectivity
  (DISFC)} and \textit{Dynamic higher-order correlations}).  We
evaluated the efficacy of this framework at recovering known patterns
in several synthetic datasets (see \textit{Synthetic data: simulating
  dynamic first-order correlations} and \textit{Synthetic data:
  simulating dynamic higher-order correlations}).  We then
applied the framework to a public fMRI dataset collected as
participants listened to an auditorily presented story, listened to a
temporally scrambled version of the story, or underwent a resting
state scan (see \textit{Functional neuroimaging data collected during
  story listening}).  We used the relative decoding accuracies of
classifiers trained on different sets of neural features to estimate
which types of features reflected ongoing cognitive processing.

\subsection*{Recovering known dynamic correlations from synthetic
  data}
\subsubsection*{Recovering dynamic first-order correlations}
We generated synthetic datasets that differed in how the underlying first-order
correlations changed over time.  For each dataset, we applied
Equation~\ref{eqn:timecorr} with a variety of kernel shapes and
widths.  We assessed how well the true underlying correlations at each
timepoint matched the recovered correlations
(Fig.~\ref{fig:synthetic}).  For every kernel and dataset we tested,
our approach recovered the correlation dynamics we embedded into the
data.  However, the quality of these recoveries varied across
different synthetic datasets in a kernel-dependent way.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/synthetic_data}
  \caption{\textbf{Recovering known dynamic first-order correlations
      from synthetic data.}  Each panel displays the average
    correlations between the vectorized upper triangles of the
    recovered correlation matrix at each timepoint and either the true
    underlying correlation at each timepoint or a reference
    correlation matrix.  (The averages are taken across 100 different
    randomly generated synthetic datasets of the given category.)
    Error ribbons denote 95\% confidence intervals (taken across
    datasets). Different colors denote different kernel shapes, and
    the shading within each color family denotes the kernel width
    parameter.  For a complete description of each synthetic dataset,
    see \textit{Synthetic data: simulating dynamic first-order
      correlations}.  \textbf{a. Constant correlations.}  These
    datasets have a stable (unchanging) underlying correlation matrix.
    \textbf{b. Random correlations.}  These datasets are generated
    using a new independently drawn correlation matrix at each new
    timepoint.  \textbf{c. Ramping correlations.}  These datasets are
    generated by smoothly varying the underlying correlations between
    the randomly drawn correlation matrices at the first and last
    timepoints.  The left panel displays the correlations between the
    recovered dynamic correlations and the underlying ground truth
    correlations.  The middle panel compares the recovered
    correlations with the \textit{first} timepoint's correlation
    matrix.  The right panel compares the recovered correlations with
    the \textit{last} timepoint's correlation matrix.
    \textbf{d. Event-based correlations.}  These datasets are each
    generated using five randomly drawn correlation matrices that each
    remain stable for a fifth of the total timecourse.  The left panel
    displays the correlations between the recovered dynamic
    correlations and the underlying ground truth correlations.  The
    right panels compare the recovered correlations with the
    correlation matrices unique to each event.  The vertical lines
    denote event boundaries.}
  \label{fig:synthetic}
\end{figure}

@@

\subsubsection*{Recovering dynamic higher-order correlations}
Following our approach to evaluating our ability to recover known
dynamic first-order correlations from synthetic data, we generated an
analogous second set of synthetic datasets that we designed to exhibit
known dynamic first-order \textit{and} second-order correlations (see
\textit{Synthetic data: simulating dynamic higher-order
  correlations}).  We generated a total of 400 datasets (100 datasets
for each category) that varied in
how the first-order and second-order correlations changed over time.
We then repeatedly applied Equation~\ref{eqn:timecorr} using the
overall best-performing kernel from our first-order tests (a Laplace
kernel with a width of 20; Fig.~\ref{fig:synthetic}) to assess how
closely the recovered dynamic correlations matched the dynamic
correlations we had embedded into the datasets.

Overall, we found that we could reliably recover both first-order and
second-order correlations from the synthetic data
(Fig.~\ref{fig:higher_order_sims}).  When the correlations were stable
for longer intervals, or changed gradually (constant, ramping, and
event datasets), recovery performance was relatively high, and we were
better able to recover dynamic first-order correlations than
second-order correlations.  This is because errors in our \textit{estimation}
procedure at lower orders necessarily propagate to higher orders
(since lower-order correlations are used to estimate higher-order
correlations).  Conversely, when the correlations were particularly
\textit{un}stable (random datasets), we better recovered second-order
correlations.  This is because noises in our \textit{data generation}
procedure propagates from higher orders to lower orders (see
\textit{Synthetic data: simulating dynamic high-order correlations}).

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/higher_order_sims}
  \caption{\textbf{Recovery of simulated first-order and second-order dynamic correlations.}  Each panel displays the average correlations
    between the vectorized upper triangles of the recovered
    first-order and second-order correlation matrices and the true
    (simulated) first-order and second order correlation matrices at each
    timepoint and for each synthetic dataset.  (The averages are taken across 100 different
    randomly generated synthetic datasets of the given category.)
    Error ribbons denote 95\% confidence intervals (taken across
    datasets).  For a complete description of each synthetic dataset,
    see \textit{Synthetic data: simulating dynamic higher-order
      correlations}.  All estimates represented in this figure were
    computed using a Laplace kernel
    (width~$=20$).  \textbf{Constant.}  These datasets
    have stable (unchanging) underlying second-order correlation
    matrices.  \textbf{Random.} These datasets are
    generated using a new independently drawn second-order correlation
    matrix at each timepoint.  \textbf{Ramping.}
    These datasets are generated by smoothly varying the underlying
    second-order correlations between the randomly drawn correlation
    matrices at the first and last timepoints.  \textbf{Event.}  These datasets are each generated using five
    randomly drawn second-order correlation matrices that each remain
    stable for a fifth of the total timecourse.  The
    vertical lines denote event boundaries.  Note that the ``dips''
    and ``ramps'' at the boundaries of sharp transitions (e.g., the
    beginning and ends of the ``constant'' and ``ramping'' datasets,
    and at the event boundaries of the ``event'' datasets) are
    finite-sample effects that reflect the reduced numbers of samples
    that may be used to accurately estimate correlations at sharp
    boundaries.}
  \label{fig:higher_order_sims}
\end{figure}

@@

\subsection*{Cognitively relevant dynamic high-order correlations in
  fMRI data}
We used across-participant temporal decoders to identify cognitively
relevant neural patterns in fMRI data (see \textit{Forward inference
  and decoding accuracy}).  The dataset we examined~\citep[collected
by][]{SimoEtal16} comprised four experimental conditions that exposed
participants to stimuli that varied systematically in how cognitively
engaging they were.  The \textit{intact} experimental condition had
participants listen to an audio recording of a 10-minute story.  The
\textit{paragraph}-scrambled experimental condition had participants
listen to a temporally scrambled version of the story, where the
paragraphs occurred out of order (but where the same total set of
paragraphs were presented over the full listening interval).  All
participants in this condition experienced the scrambled paragraphs in
the same order.  The \textit{word}-scrambled experimental condition
had participants listen to a temporally scrambled version of the story
where the words in the story occurred in a random order.  All
participants in the word condition experienced the scrambled words in
the same order.  Finally, in a \textit{rest} experimental condition,
participants lay in the scanner with no overt stimulus, with their
eyes open (blinking as needed).  This public dataset provided a convenient
means of testing our hypothesis that different levels of cognitive
processing and engagement are reflected in different orders of brain
activity dynamics.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/decode_level}
  \caption{\textbf{Across-participant timepoint decoding accuracy varies with
      correlation order and cognitive engagement.}
    \textbf{a.~Decoding accuracy as a function of order: PCA.}
    \textit{Order} ($x$-axis) refers to the maximum order of dynamic
    correlations that were available to the classifiers (see
    \textit{Feature weighting and testing}).  The reported
    across-participant decoding accuracies are averaged over all
    kernel shapes and widths (see \textit{Identifying robust decoding
      results}).  The $y$-values are displayed relative to chance
    accuracy (intact: $\frac{1}{300}$; paragraph: $\frac{1}{272}$;
    word: $\frac{1}{300}$; rest: $\frac{1}{400}$; these chance accuracies
    were subtracted from the observed accuracies to obtain the
    relative accuracies reported on the $y$-axis).  The error ribbons
    denote 95\% confidence intervals across cross-validation folds
    (i.e., random assignments of participants to the training and test
    sets).  The colors denote the experimental condition.  Arrows
    denote sets of features that yielded reliably higher (upward
    facing) or lower (downward facing) decoding accuracy than the mean
    of all other features (via a two-tailed $t$-test, thresholded at
    $p < 0.05$).  Figure~\ref{fig:ttests} displays additional
    comparisons between the decoding accuracies achieved using
    different sets of neural features.  The circled values represent
    the maximum decoding accuracy within each experimental condition.
    \textbf{b.~Normalized timepoint decoding accuracy as a function of order:
      PCA.}  This panel displays the same results as Panel a, but here
    each curve has been normalized to have a maximum value of 1 and a
    minimum value of 0 (including the upper and lower bounds of the
    respective 95\% confidence intervals).  Panels a and b used PCA to
    project each high-dimensional pattern of dynamic correlations onto
    a lower-dimensional space.  \textbf{c.~Timepoint decoding accuracy as a
      function of order: eigenvector centrality.} This panel is in the
    same format as Panel a, but here eigenvector centrality has been
    used to project the high-dimensional patterns of dynamic
    correlations onto a lower-dimensional space.
    \textbf{d.~Normalized timepoint decoding accuracy as a function of order:
      eigenvector centrality.} This panel is in the same format as
    Panel b, but here eigenvector centrality has been used to project
    the high-dimensional patterns of dynamic correlations onto a
    lower-dimensional space. See Figures~\kernelshape~and~\kernelwidth~for decoding results broken down by kernel shape and width, respectively.}
  \label{fig:decoding}
\end{figure}

In brief, we computed timeseries of dynamic high-order correlations
that were similar across participants in each of two randomly assigned
groups: a training group and a test group.  We then trained
classifiers on the training group's data to match each sample from the
test group with a stimulus timepoint.  Each classifier comprised a
weighted blend of neural patterns that reflected up to
$n^\mathrm{th}$-order dynamic correlations (see \textit{Feature
  weighting and testing}, Fig.~\ref{fig:pipeline}).  We repeated this process for
$n \in \left\{ 0, 1, 2, ..., 10 \right\}$.  Our examinations of
synthetic data suggested that none of the kernels we examined were
``universal'' in the sense of optimally recovering underlying
correlations regardless of the temporal structure of those
correlations.  We found a similar pattern in the (real) fMRI data,
whereby different kernels yielded different decoding accuracies, but
no single kernel emerged as the clear ``best.''  In our analyses of
neural data, we therefore averaged our decoding results over a variety
of kernel shapes and widths in order to identify results that were
robust to specific kernel parameters (see \textit{Identifying robust
  decoding results}).

@@

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/stats_heatmaps}
\caption{\textbf{Statistical summary of decoding accuracies for
    different neural features.}  Each column of matrices displays decoding
  results for one experimental condition (intact, paragraph, word, and
  rest).  We considered dynamic activity patterns (order 0) and
  dynamic correlations at different orders (order $> 0$).  We used
  two-tailed $t$-tests to compare the
  distributions of decoding accuracies obtained using each pair of
  features.  The distributions for each feature reflect the set of
  average decoding accuracies (across all kernel parameters), obtained
  for each random assignment of training and test groups.
  In the upper triangles of each matrix, warmer colors (positive $t$-values) indicate that the
  neural feature indicated in the given row yielded higher accuracy than the
  feature indicated in the given column.  Cooler colors (negative
  $t$-values) indicate that
  the feature in the given row yielded lower decoding accuracy than
  the feature in the given column.  The lower triangles of each map
  denote the corresponding $p$-values for the $t$-tests.  The diagonal
  entries display the relative average optimized weight given to each type of feature in
  a decoder that included all feature types (see \textit{Feature
    weighting and testing}).}
\label{fig:ttests}
\end{figure}

@@

\bibliographystyle{apacite}
\bibliography{CDL-bibliography/cdl}

>>>>

\documentclass[english]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage[sort&compress, numbers, super]{natbib}
\usepackage{pxfonts}
\usepackage[utf8]{inputenc}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage[left]{lineno}
\usepackage{soul}
\usepackage[nofiglist, fighead]{endfloat}
\linenumbers

@@

\begin{abstract}
  Our thoughts arise from coordinated patterns of interactions between
  brain structures that change with our ongoing experiences.
  High-order dynamic correlations in neural activity patterns reflect
  different subgraphs of the brain's functional connectome that display
  homologous lower-level dynamic correlations.  Here we test the
  hypothesis that high-level cognition is reflected in high-order
  dynamic correlations in brain activity patterns.  We develop an
  approach to estimating high-order dynamic correlations in timeseries
  data, and we apply the approach to neuroimaging data collected as
  human participants either listen to a ten-minute story or listen
  to a temporally scrambled version of the story.  We train
  across-participant pattern classifiers to decode (in held-out data)
  when in the session each neural activity snapshot was collected.  We
  find that classifiers trained to decode from high-order dynamic
  correlations yield the best performance on data collected as
  participants listened to the (unscrambled) story.  By contrast,
  classifiers trained to decode data from scrambled versions of the
  story yielded the best performance when they were trained using
  first-order dynamic correlations or non-correlational activity
  patterns.  We suggest that as our thoughts become more complex, they
  are reflected in higher-order patterns of dynamic network
  interactions throughout the brain.
\end{abstract}

@@

\section*{Introduction}
A central goal in cognitive neuroscience is to elucidate the
neural code: the mapping between (a) mental states or
cognitive representations and (b) neural activity patterns. One means
of testing models of the neural code is to ask how accurately that
model is able to ``translate'' neural activity patterns into known (or
hypothesized) mental states or cognitive
representations~\citep[e.g.,][]{HaxbEtal01, NormEtal06b, TongPrat12,
  MitcEtal08a, KamiTong05, NishEtal11, PereEtal18, HuthEtal12,
  HuthEtal16}.  Training decoding models on different types of neural
features (Fig.~\ref{fig:patterns}a) can also help to elucidate which
specific aspects of neural activity patterns are informative about
cognition and, by extension, which types of neural activity patterns
might compose the neural code.  For example, prior work has used
region of interest analyses to estimate the anatomical locations of
specific neural representations~\citep[e.g.,][]{EtzeEtal09}, or to
compare the relative contributions to the neural code of multivariate
activity patterns versus dynamic correlations between neural activity
patterns~\citep[e.g.,][]{MannEtal18, FongEtal19}.  An emerging theme
in this literature is that cognition is mediated by dynamic
interactions between brain structures~\citep{Gros88, Fris00, SporHone06, BassEtal06,
  Turk13, DemeEtal19, SoloEtal19, LuriEtal18, PretEtal17, ZouEtal19,
  MackEtal17, BresKels01, McIn00}.

@@

Studies of the neural code to date have primarily focused on
univariate or multivariate neural patterns~\citep[for review
see][]{NormEtal06b}, or (more recently) on patterns of dynamic
first-order correlations~\citep[i.e., interactions between pairs of
brain structures;][]{MannEtal18, FongEtal19, LuriEtal18, PretEtal17,
  ZouEtal19, DemeEtal19}.  What might the future of this line of work
hold?  For example, is the neural code implemented through
higher-order interactions between brain structures~\citep[e.g.,
see][]{ReimEtal17}?  Second-order correlations reflect
homologous patterns of correlation.  In other words, if the
dynamic patterns of correlations between two regions, $A$ and $B$, are
similar to those between two other regions, $C$ and $D$, this would be
reflected in the second-order correlations between ($A$--$B$) and
($C$--$D$).  In this way, second-order correlations identify
similarities and differences between subgraphs of the brain's
connectome.  Analogously, third-order correlations reflect homologies
between second-order correlations-- i.e., homologous patterns of
homologous interactions between brain regions.  More generally,
higher-order correlations reflect homologies between patterns of
lower-order correlations.  We can then ask: which ``orders'' of
interaction are most reflective of high-level cognitive processes?

@@

To gain insights into the above aspects of the neural code, we
developed a computational framework for estimating dynamic high-order
correlations in timeseries data. This framework provides an important
advance, in that it enables us to examine patterns of higher-order
correlations that are computationally intractable to estimate via
conventional methods.  Given a multivariate timeseries, our framework
provides timepoint-by-timepoint estimates of the first-order
correlations, second-order correlations, and so on.  Our approach
combines a kernel-based method for computing dynamic correlations in
timeseries data with a dimensionality reduction step
(Fig.~\ref{fig:patterns}b) that projects the resulting dynamic
correlations into a low-dimensional space.  We explored two
dimensionality reduction approaches: principle components
analysis~\citep[PCA;][]{Pear01}, which preserves an approximately
invertible transformation back to the original data~\citep[e.g., this
follows related approaches taken by][]{McInJirs19, TokeSomm19,
  GonzEtal19}; and a second non-invertible algorithm for computing
dynamic patterns in eigenvector centrality~\citep{Land95}.  This
latter approach characterizes correlations between each feature
dimension's relative position in the network (at each moment
in time) in favor of the specific activity histories of different
features~\citep[also see][]{BetzEtal19, SizeEtal18, ReimEtal17}.

@@

\section*{Results}
We sought to understand whether high-level cognition is reflected in
dynamic patterns of high-order correlations.  To that end, we
developed a computational framework for estimating the dynamics of
stimulus-driven high-order correlations in multivariate timeseries
data (see Dynamic inter-subject functional connectivity
  (DISFC) and Dynamic higher-order correlations).  We
evaluated the efficacy of this framework at recovering known patterns
in several synthetic datasets (see Synthetic data: simulating
  dynamic first-order correlations and Synthetic data:
  simulating dynamic higher-order correlations).  We then
applied the framework to a public fMRI dataset collected as
participants listened to an auditorily presented story, listened to a
temporally scrambled version of the story, or underwent a resting
state scan (see Functional neuroimaging data collected during
  story listening).  We used the relative decoding accuracies of
classifiers trained on different sets of neural features to estimate
which types of features reflected ongoing cognitive processing.

\subsection*{Recovering dynamic first-order correlations}
We generated synthetic datasets that differed in how the underlying first-order
correlations changed over time.  For each dataset, we applied
Equation~\ref{eqn:timecorr} with a variety of kernel shapes and
widths.  We assessed how well the true underlying correlations at each
timepoint matched the recovered correlations
(Fig.~\ref{fig:synthetic}).  For every kernel and dataset we tested,
our approach recovered the correlation dynamics we embedded into the
data.  However, the quality of these recoveries varied across
different synthetic datasets in a kernel-dependent way.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/synthetic_data}
  \caption{\textbf{Recovering known dynamic first-order correlations
      from synthetic data.}  Each panel displays the average
    correlations between the vectorized upper triangles of the
    recovered correlation matrix at each timepoint and either the true
    underlying correlation at each timepoint or a reference
    correlation matrix.  (The averages are taken across 100 different
    randomly generated synthetic datasets of the given category.)
    Error ribbons denote 95\% confidence intervals (taken across
    datasets). Different colors denote different kernel shapes, and
    the shading within each color family denotes the kernel width
    parameter.  For a complete description of each synthetic dataset,
    see Synthetic data: simulating dynamic first-order
      correlations.  \textbf{a. Constant correlations.}  These
    datasets have a stable (unchanging) underlying correlation matrix.
    \textbf{b. Random correlations.}  These datasets are generated
    using a new independently drawn correlation matrix at each new
    timepoint.  \textbf{c. Ramping correlations.}  These datasets are
    generated by smoothly varying the underlying correlations between
    the randomly drawn correlation matrices at the first and last
    timepoints.  The left panel displays the correlations between the
    recovered dynamic correlations and the underlying ground truth
    correlations.  The middle panel compares the recovered
    correlations with the first timepoint's correlation
    matrix.  The right panel compares the recovered correlations with
    the last timepoint's correlation matrix.
    \textbf{d. Event-based correlations.}  These datasets are each
    generated using five randomly drawn correlation matrices that each
    remain stable for a fifth of the total timecourse.  The left panel
    displays the correlations between the recovered dynamic
    correlations and the underlying ground truth correlations.  The
    right panels compare the recovered correlations with the
    correlation matrices unique to each event.  The vertical lines
    denote event boundaries.}
  \label{fig:synthetic}
\end{figure}

@@

\subsection*{Recovering dynamic higher-order correlations}
Following our approach to evaluating our ability to recover known
dynamic first-order correlations from synthetic data, we generated an
analogous second set of synthetic datasets that we designed to exhibit
known dynamic first-order and second-order correlations (see
Synthetic data: simulating dynamic higher-order
  correlations).  We generated a total of 400 datasets (100 datasets
for each category) that varied in
how the first-order and second-order correlations changed over time.
We then repeatedly applied Equation~\ref{eqn:timecorr} using the
overall best-performing kernel from our first-order tests (a Laplace
kernel with a width of 20; Fig.~\ref{fig:synthetic}) to assess how
closely the recovered dynamic correlations matched the dynamic
correlations we had embedded into the datasets.

Overall, we found that we could reliably recover both first-order and
second-order correlations from the synthetic data
(Fig.~\ref{fig:higher_order_sims}).  When the correlations were stable
for longer intervals, or changed gradually (constant, ramping, and
event datasets), recovery performance was relatively high, and we were
better able to recover dynamic first-order correlations than
second-order correlations.  This is because errors in our estimation
procedure at lower orders necessarily propagate to higher orders
(since lower-order correlations are used to estimate higher-order
correlations).  Conversely, when the correlations were particularly
unstable (random datasets), we better recovered second-order
correlations.  This is because noises in our data generation
procedure propagates from higher orders to lower orders (see
Synthetic data: simulating dynamic high-order correlations).

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/higher_order_sims}
  \caption{\textbf{Recovery of simulated first-order and second-order dynamic correlations.}  Each panel displays the average correlations
    between the vectorized upper triangles of the recovered
    first-order and second-order correlation matrices and the true
    (simulated) first-order and second order correlation matrices at each
    timepoint and for each synthetic dataset.  (The averages are taken across 100 different
    randomly generated synthetic datasets of the given category.)
    Error ribbons denote 95\% confidence intervals (taken across
    datasets).  For a complete description of each synthetic dataset,
    see Synthetic data: simulating dynamic higher-order
      correlations.  All estimates represented in this figure were
    computed using a Laplace kernel
    (width~$=20$).  \textbf{Constant.}  These datasets
    have stable (unchanging) underlying second-order correlation
    matrices.  \textbf{Random.} These datasets are
    generated using a new independently drawn second-order correlation
    matrix at each timepoint.  \textbf{Ramping.}
    These datasets are generated by smoothly varying the underlying
    second-order correlations between the randomly drawn correlation
    matrices at the first and last timepoints.  \textbf{Event.}  These datasets are each generated using five
    randomly drawn second-order correlation matrices that each remain
    stable for a fifth of the total timecourse.  The
    vertical lines denote event boundaries.  Note that the ``dips''
    and ``ramps'' at the boundaries of sharp transitions (e.g., the
    beginning and ends of the ``constant'' and ``ramping'' datasets,
    and at the event boundaries of the ``event'' datasets) are
    finite-sample effects that reflect the reduced numbers of samples
    that may be used to accurately estimate correlations at sharp
    boundaries.}
  \label{fig:higher_order_sims}
\end{figure}

@@

\subsection*{Cognitively relevant dynamic high-order correlations in
  fMRI data}
We used across-participant temporal decoders to identify cognitively
relevant neural patterns in fMRI data (see Forward inference
  and decoding accuracy).  The dataset we examined~\citep[collected
by][]{SimoEtal16} comprised four experimental conditions that exposed
participants to stimuli that varied systematically in how cognitively
engaging they were.  The intact experimental condition had
participants listen to an audio recording of a 10-minute story.  The
paragraph-scrambled experimental condition had participants
listen to a temporally scrambled version of the story, where the
paragraphs occurred out of order (but where the same total set of
paragraphs were presented over the full listening interval).  All
participants in this condition experienced the scrambled paragraphs in
the same order.  The word-scrambled experimental condition
had participants listen to a temporally scrambled version of the story
where the words in the story occurred in a random order.  All
participants in the word condition experienced the scrambled words in
the same order.  Finally, in a rest experimental condition,
participants lay in the scanner with no overt stimulus, with their
eyes open (blinking as needed).  This public dataset provided a convenient
means of testing our hypothesis that different levels of cognitive
processing and engagement are reflected in different orders of brain
activity dynamics.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/decode_level}
  \caption{\textbf{Across-participant timepoint decoding accuracy varies with
      correlation order and cognitive engagement.}
    \textbf{a.~Decoding accuracy as a function of order: PCA.}
    Order ($x$-axis) refers to the maximum order of dynamic
    correlations that were available to the classifiers (see
    Feature weighting and testing).  The reported
    across-participant decoding accuracies are averaged over all
    kernel shapes and widths (see Identifying robust decoding
      results).  The $y$-values are displayed relative to chance
    accuracy (intact: $\frac{1}{300}$; paragraph: $\frac{1}{272}$;
    word: $\frac{1}{300}$; rest: $\frac{1}{400}$; these chance accuracies
    were subtracted from the observed accuracies to obtain the
    relative accuracies reported on the $y$-axis).  The error ribbons
    denote 95\% confidence intervals across cross-validation folds
    (i.e., random assignments of participants to the training and test
    sets).  The colors denote the experimental condition.  Arrows
    denote sets of features that yielded reliably higher (upward
    facing) or lower (downward facing) decoding accuracy than the mean
    of all other features (via a two-tailed $t$-test, thresholded at
    $p < 0.05$).  Figure~\ref{fig:ttests} displays additional
    comparisons between the decoding accuracies achieved using
    different sets of neural features.  The circled values represent
    the maximum decoding accuracy within each experimental condition.
    \textbf{b.~Normalized timepoint decoding accuracy as a function of order:
      PCA.}  This panel displays the same results as Panel a, but here
    each curve has been normalized to have a maximum value of 1 and a
    minimum value of 0 (including the upper and lower bounds of the
    respective 95\% confidence intervals).  Panels a and b used PCA to
    project each high-dimensional pattern of dynamic correlations onto
    a lower-dimensional space.  \textbf{c.~Timepoint decoding accuracy as a
      function of order: eigenvector centrality.} This panel is in the
    same format as Panel a, but here eigenvector centrality has been
    used to project the high-dimensional patterns of dynamic
    correlations onto a lower-dimensional space.
    \textbf{d.~Normalized timepoint decoding accuracy as a function of order:
      eigenvector centrality.} This panel is in the same format as
    Panel b, but here eigenvector centrality has been used to project
    the high-dimensional patterns of dynamic correlations onto a
    lower-dimensional space. See Figures~\kernelshape~and~\kernelwidth~for decoding results broken down by kernel shape and width, respectively.}
  \label{fig:decoding}
\end{figure}

In brief, we computed timeseries of dynamic high-order correlations
that were similar across participants in each of two randomly assigned
groups: a training group and a test group.  We then trained
classifiers on the training group's data to match each sample from the
test group with a stimulus timepoint.  Each classifier comprised a
weighted blend of neural patterns that reflected up to
$n^\mathrm{th}$-order dynamic correlations (see Feature
  weighting and testing, Fig.~\ref{fig:pipeline}).  We repeated this process for
$n \in \left\{ 0, 1, 2, ..., 10 \right\}$.  Our examinations of
synthetic data suggested that none of the kernels we examined were
``universal'' in the sense of optimally recovering underlying
correlations regardless of the temporal structure of those
correlations.  We found a similar pattern in the (real) fMRI data,
whereby different kernels yielded different decoding accuracies, but
no single kernel emerged as the clear ``best.''  In our analyses of
neural data, we therefore averaged our decoding results over a variety
of kernel shapes and widths in order to identify results that were
robust to specific kernel parameters (see Identifying robust
  decoding results).

@@

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/stats_heatmaps}
\caption{\textbf{Statistical summary of decoding accuracies for
    different neural features.}  Each column of matrices displays decoding
  results for one experimental condition (intact, paragraph, word, and
  rest).  We considered dynamic activity patterns (order 0) and
  dynamic correlations at different orders (order $> 0$).  We used
  two-tailed $t$-tests to compare the
  distributions of decoding accuracies obtained using each pair of
  features.  The distributions for each feature reflect the set of
  average decoding accuracies (across all kernel parameters), obtained
  for each random assignment of training and test groups.
  In the upper triangles of each matrix, warmer colors (positive $t$-values) indicate that the
  neural feature indicated in the given row yielded higher accuracy than the
  feature indicated in the given column.  Cooler colors (negative
  $t$-values) indicate that
  the feature in the given row yielded lower decoding accuracy than
  the feature in the given column.  The lower triangles of each map
  denote the corresponding $p$-values for the $t$-tests.  The diagonal
  entries display the relative average optimized weight given to each type of feature in
  a decoder that included all feature types (see Feature
    weighting and testing).}
\label{fig:ttests}
\end{figure}

@@

\section*{Competing interests}
The authors declare no competing financial interests.

\bibliographystyle{naturemag}
\bibliography{CDL-bibliography/cdl}
